Drought indices of water demand and supply and their impact on LULCC in Chile from ERA5-Land and MODIS

### 2024-01-26 article/paper_drought_Chile2023.tex
actualizar repo

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

In contrast to our observations, \citep{Venegas2022}, who investigated
Cryptocarya alba and Beilschmiedia miersii (both members of the
Lauraceae family) inhabiting sclerophyllous forests in Chile, concluded
that the overall development of the trees had decelerated, potentially
indicating alterations in the inherent dynamics of their respective
forests. They attributed it to the cumulative effects of the
unprecedented drought (i.e., hydrological drought). Nonetheless, the
trees within sclerophyllous forests are highly heterogeneous, and the
species analyzed by \citep{Venegas2022} are not the predominant ones.
Because of this, when we used MOD13A3 with a 1km spatial resolution to
measure vegetation condition, it took the average condition of 1 square
kilometer. Thus, we attribute that forest to being the most resistant to
drought, due to the fact that most of the species comprising it are
highly resilient to water scarcity. Then, for a study of the dynamics of
forest to drought, e.g., sclerophyllous forest, on a local scale using
remote sensing, we should use operational products at higher spatial
resolutions as provided by the missions Landsat or Sentinel that could
capture the heterogenity of the forest type.

@@

The main external factors that affect biomass production by vegetation
are ET and SM, and the rate of ET in turn depends on the availability of
water storage in the root zone. Thus, soil moisture plays a key role in
land carbon uptake and, consequently, in the production of biomass
\citep{Humphrey2021}. Moreover, \citep{Zhang2022} indicates there is a
bidirectional causality between soil moisture and vegetation
productivity. For this, lastly, are some studies that have redefined
agricultural drought as soil moisture drought from the hydrological
perspective \citep{Loon2016, Samaniego2018}. Even though soil moisture
is the external factor most determinant of vegetation biomass, there are
multiple internal factors, such as species, physiological
characteristics, and plant hydraulics, that would affect vegetation
productivity. Because of that, we believe that agricultural drought,
referring to the drought that impacts vegetation productivity, is the
most proper term, as originally defined by \citep{Wilhite1985}.

Our results showed that the drought index derived from soil moisture
(SSI) was the most effective in explaining vegetation productivity
within landcover macroclasses. However, there is strong spatial
variability throughout Chile, mainly attributable to the climate
heterogeneity.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

In contrast to our observations, \citep{Venegas2022}, who investigated
Cryptocarya alba and Beilschmiedia miersii (both members of the
Lauraceae family) inhabiting sclerophyllous forests in Chile, concluded
that the overall development of the trees had decelerated, potentially
indicating alterations in the inherent dynamics of their respective
forests. They attributed it to the cumulative effects of the
unprecedented drought (i.e., hydrological drought). Thus, we attribute
that forest to being the most resistant to drought, due to the fact that
most of the species comprising it are highly resilient to water
scarcity. compared to the other landcover classes. Also, when we used
MOD13A3 with a 1km spatial resolution to measure vegetation condition,
it took the average condition of 1 square kilometer. Then, to study how
a type of forest (sclerophyllous forest) changes in response to drought
on a local level using remote sensing, we should use operational
products with higher spatial resolutions, such as those from Landsat or
Sentinel, which can show how different the forests are.

@@

The main external factors that affect biomass production by vegetation
are ET and SM, and the rate of ET in turn depends on the availability of
water storage in the root zone. Thus, soil moisture plays a key role in
land carbon uptake and, consequently, in the production of biomass
\citep{Humphrey2021}. Moreover, \citep{Zhang2022} indicates there is a
bidirectional causality between soil moisture and vegetation
productivity. For this, lastly, some studies have redefined agricultural
drought as soil moisture drought from the hydrological perspective
\citep{Loon2016, Samaniego2018}. Even though soil moisture is the
external factor most determinant of vegetation biomass, there are
multiple internal factors, such as species, physiological
characteristics, and plant hydraulics, that would affect vegetation
productivity. Because of that, we believe that agricultural drought,
referring to the drought that impacts vegetation productivity, is the
most proper term, as originally defined by \citep{Wilhite1985}.

Our results showed that the drought index derived from soil moisture
(SSI) was the most effective in explaining vegetation productivity
within landcover macroclasses. However, there is strong spatial
variability throughout Chile and between classes, mainly attributable to
climate heterogeneity, hydrological status, or vegetation resistance to
water scarcity. The semi-arid ``Norte Chico'' and the Mediterranean
``Centro'' were where SSI had the best performance. Medium-term deficits
of 12 months are more relevant in the response of vegetation, which
decreases to the south, and in the case of croplands, they seem to react
in a shorter time, with six months (SSI-6) in ``Centro.''
